\input{preamble}

\begin{document}

\title{Deformation theory}
\maketitle

\phantomsection
\label{section-phantom}
\hfill
\href{http://github.com/danimalabares/stack}{github.com/danimalabares/stack}

\tableofcontents

\section{Introduction}
\label{section-intro}

\noindent
\cite{Sernesi-Overview} has a great way of motivating deformation
theory via the moduli problem.
Here let's just put the following important meaning for
``deformation theory'' (as opposed to studying flat families!):

\begin{quote}
``{\it Deformation theory} is the study of infinitesimal deformaitons
as a tool to understand the local structure of the moduli space.
The goal is to be able to describe the restriction
of the universal family to a small neighborhood of $m \in \mathcal{M}$,
or, more precisely, its restriction to the germ of $M$ at $m$.

What is interesting here is that we can study order and infinitesimal
deformations even though the functor $F$ is not representable or simply we don’t
yet know it is. This is the most frequent case. Such an investigation will
reveal the infinitesimal properties at $[X]$ of a yet unknown global structure
on $M$ which will be hopefully understood at a subsequent stage of the
investigation.  In order words it turns out to be possible and convenient to
separate the {\it global moduli problem} from the {\it local moduli problem},
and deformation theory studies the latter, with the purpose of constructing a
family of deformations of a given object parametrized by the spectrum of a local
ring, and having properties as close as possible to a universal property.''
\end{quote}



\section{Extensions}
\label{section-extensions}

\noindent
In this section I discuss extensions of rings
and their geometric counterpart.

\begin{definition}
\label{definition-extension}
\begin{reference}
\cite[p.7]{def0}
\end{reference}
Let $A \to R$ be a ring homomorphism
(i.e. $R$ is an $A$-algebra).
An {\it $A$-extension of $R$ by $I$} is a short exact sequence
$$
\xymatrix{
0\ar[r]
&I\ar[r]
&E\ar[r]^\varphi
&R\ar[r]
&0.
}
$$
where $E$ is an $R'$-algebra
and $\varphi$ is a morphism of $A$-algebras
whose kernel is $I$.
We also require that $I^2=(0)$;
this condition implies that $I$ has the sturcture of $R$-module.
\end{definition}

\begin{example}
\label{example-dual-numbers-extension}
$$
\xymatrix{
0\ar[r]
&(\varepsilon)\ar[r]
&A[\varepsilon]\ar[r]
&A\ar[r]
&0
}
$$
where $A[\varepsilon]$ is the algebra of dual numbers over $A$.
\end{example}

\noindent
Without explaining details,
let me for now just say that there are notions of
morphism of extensions, pushforward, pullbacks,
and even a sum with makes the category of
extensions of $A$-algebras $R$ by $I$, denoted by
$\text{Ex}_A(R,I)$, an $R$-module. More precisely:

\begin{proposition}
\label{proposition-extensions-are-module}
\cite[Proposition 1.1.4]{def0}.
\end{proposition}

\noindent
Next in line: define the first cotangent module
(so that I may understand the grading of the first
cotangent module of a SR ring…)


\section{Deformations of schemes}
\label{section-deformations-of-schemes}

\noindent
The following definition is from [lucas-defos], which in turn comes from
\cite{Sernesi-deformations}

\begin{definition}
\label{definition-deformation}
Let $X$ be an algebraic $\mathbb{C}$-scheme.
\begin{enumerate}
\item A {\it deformation} of $X$ is a Cartesian diagram $\xi$
$$
\xymatrix{
X\ar[r]\ar[d]&\mathcal{X}\ar[d]^{\pi}\\
\Spec \mathbb{C}\ar[r]^s&S
}
$$
where $\pi$ is a flat surjective morphism of algebraic $\mathbb{C}$-schemes and
$S$ is connected. (Recall that flatness accounts for ``continuity''.)
\item A {\it local deformation} of $X$ is a deformation $\xi$ where
$S=\Spec A$ for $A$ a noetherian local $\mathbb{C}$ algebra with residue
field $\mathbb{C}$.
\item An {\it infinitesimal deformation of $X$} is a local deformation with $A$
an artinian local $\mathbb{C}$-algebra with residue field $\mathbb{C}$. $X$ is
called {\it rigid} if all infinitesimal deformations are trivial.
\item An {\it inifinitesimal deformation of order $n$} is an infinitesimal
deformation when $S=\Spec (\mathbb{C}[t]/(t^{n+1})$.
\end{enumerate}
\end{definition}

\section{Kodaira-Spencer map}
\label{section-kodaira-spencer-map}


\section{Kodaira-Spencer correspondence}
\label{section-kodaia-spencer-correspondence}


\medskip\noindent
After a few months of the previous text in this section
I think I have understood the construction of
Kodaira-Spencer map for embedded deformations,
which is \cite[Proposition 3.2.1(ii)]{Sernesi-deformations}.

But I will actually briefly outline a particular case
of this, which is easier: \cite[Examples 3.2.4(ii)]{Sernesi-deformations}.
This the case when $X \subset Y$ is a Cartier divisor.
Then $X$ is cut out locally by a set of functions $f_i$
defined on an affine cover $U_i$ of $Y$.
Then we have a line bundle defined by the
transition functions $f_{ij}=f_i/f_j$.
This is the line bundle $\mathcal{O}_Y(X)$, which
by the adjunction formula is also the normal bundle $N_{X/Y}$
of $X$ in $Y$.

Now consider a first order deformation
$$
\xymatrix{
X\ar[r]\ar[d]
&\mathcal{X}\subset Y\times B_\varepsilon\ar[d]^\pi\\
0\ar[r]
&B_\varepsilon
}
$$

The great thing about first order deformations
is that the structure sheaf of $\mathcal{X}$
on an affine open $U_i$ of $Y$ 
looks like the tensor product 
$\mathcal{O}_{U_i}\otimes \mathbb{C}[\varepsilon]/(\varepsilon^2)$
(\href{https://stacks.math.columbia.edu/tag/01I4}{01I4}).
By definition of tensor product this just says
that sections look like $f+\varepsilon g$
for $f,g \in \Gamma(\mathcal{O}_{U_i})$.

If $\mathcal{X}$ is a Cartier divisor on $Y \times B_\varepsilon$
(it is; but I'm missing a formal proof;
the intuition is that an ``infinitesimal thickening'' $\mathcal{X}$ 
of $X$ has the same underlying topological space
but different scheme structure; 
\href{https://stacks.math.columbia.edu/tag/01JL}{01JL}
\href{https://stacks.math.columbia.edu/tag/01I4}{01I4}),
it is given as the vanishing locus
of some local sections $f_i+\varepsilon g_i$.
(Perhaps the $f_i$ coincide with the distinguished
section of $X$.)
The line bundle associated to $\mathcal{X}$ 
is given by the cocycle
$$
F_{ij}=f_{ij}+\varepsilon g_{ij}
$$
for $g_{ij}=\frac{g_i-f_{ij}g_j}{f_j}$.
Indeed, note that 
$\frac{1}{f_j+\varepsilon g_j}=\frac{1}{f_j}-\varepsilon \frac{g_j}{f^2_j}$ 
and compute $F_{ij}=(f_i+\varepsilon g_j)/(f_j+\varepsilon g_j)$.

Then
$$
f_i+\varepsilon g_i=(f_{ij}+\varepsilon g_{ij})(f_j+\varepsilon g_j)
$$
gives
$$
g_i=f_{ij}g_j+g_{ij}f_j
$$
and since $f_j$ vanishes along $X$,
we are left with $g_i=f_{ij}g_j$ 
which is how a section of the sheaf $N_{X/Y}$
is constructed by definition of sheaf.


\begin{example}
\label{example-moduli-space-of-nonsingular-Riemann-surfaces-of-genus-g}
\begin{reference}
\cite[Example 3.1]{continued-fractions}
\end{reference}
Fix $g \geq 2$. The dimension of the deformation space of a nonsingular
projective curve $X$ is $3g-3$. This is ``the dimension of the moduli space of
curves of genus $g$''.
 
We can compute this number by Riemann-Roch formula on the bundle $-K_X$. 
Indeed, since $X$ is a curve, $\Omega_X^1=K_X$ and thus $-K_X=T_X$. We get
\begin{align*}
h^0(-K)-h^0(K-(-K))&=\text{deg}(-K)-g+1\\
&=-2g+2-g+1\qquad \substack{\text{degree additive and}\\ \text{deg}K=2g-2}\\
&=-3g+3
\end{align*}
Now by Serre duality $h^0(K-(-K))=h^1(-K)=h^1(T_X)$, and it turns out that a
Riemann surface of genus $g \geq 2$ has no holomorphic vector fields, so that
$h^0(-K)=h^0(T_X)=0$.
\end{example}

\begin{exercise}
\label{exercise-deformations-of-curves-in-K3}
Let $C$ be a smooth genus $g$ curve which can be embedded in a K3 surface  $M$,
and $X$ the family of all deformations of $C$ in $M$.
\begin{enumerate}
\item Prove that $\dim X \leq  g$.
\item Let $\mathcal{X}_g$ be the space of all curves of genus $g$ (smooth?)
which can be possibly embedded to a K3 surface. Prove that each irreducible
component $Z$ of $\mathcal{X}_g$ satisfies $\dim_\mathbb{C} \leq  g+19$. Deduce
that there exists a compact complex curve which cannot be embedded in a K3
surface.
\end{enumerate}
\end{exercise}

\begin{exercise}
\label{exercise-}
Let $C$ be a smooth curve embedded in a K3 surface $X$. Show that the dimension
of the deformation space of $C$ is $\leq g$.
\end{exercise}

\begin{proof}
%\begin{enumerate}
%\item 
The deformation space of a variety is the space of isomorphism classes of
deformations as explained above. It turns out that there is a way to associate
1-cocyles of the tangent sheaf to deformations, so that in fact the deformation
space $\text{Def}_1$ is isomorphic to $H^{1}(X,\mathcal{T}_X)$ for any variety
$X$.

For our curve $C$ we thus know that the dimension of the space of deformations
(deformations not necessarily contained in $X$) is $h^1(\mathcal{T}_C)$.  The
family of deformations of $C$ that are contained in $X$ is the Hilbert scheme of
curves with fixed Hilbert polynomial $P(t)$ after quotienting by $\mathbb{C}s$,
where $s$ is the section whose vanishing locus is  $C$. 
This says that the number we are looking for is $h^{0}(X,\mathcal{O}(C))-1$.

Before showing that in fact $h^0(X,\mathcal{O}(C))=1+g$,
let's explain why this Hilbert scheme construction actually works.
Intuitively, any curve lying in the same component
as $C$ would be deformation-equivalent to $C$
--- at least we know that it's Hilbert polynomial is the same.
But it's not clear what's the deformation involving these two
(this question should be addressed!).
Further, this doesn't explain Vladimiro's
suggesion of ``taking quotient by $\mathbb{C}s$''.
To me it looks like I would just compute the dimension
of the tangent space of $\text{Hilb}^P(M)$ at $C$.

Now I will show that $h^0(X,\mathcal{O}(C))=1+g$ (see 
\cite[Lemma 1.2.1, Remark 1.2.2]{huk}).

Consider the ideal sheaf exact sequence twisted by
$\mathcal{O}(C)$:
$$
\xymatrix{
0\ar[r]&\mathcal{O}_X\ar[r]&\mathcal{O}(C)\ar[r]
&\mathcal{O}_X(C) \otimes \mathcal{O}_C=\mathcal{O}(C)|_{C}\ar[r]&0
}
$$
By $X$ being a K3 we know that $H^{1}(\mathcal{O}_X)=0$, so that we have the
short exact sequence in cohomology
$$
\xymatrix{
0\ar[r]&H^{0}(\mathcal{O}_X)\ar[r]&H^{0}(\mathcal{O}(C))\ar[r]
&H^{0}(\mathcal{O}(C)|_{C})\ar[r]&0
}
$$
so that $h^0(\mathcal{O}(C))=1+h^1(\mathcal{O}(C)|_{C})$. 
A version of adjunction formula says
$\omega_C \cong (K_X \otimes\mathcal{O}(C)|_{C}$, and using that 
$K_X=\mathcal{O}_X$ we obtain $h^0(\mathcal{O}(C)|_{C})=h^0(\omega_C)=g$.

To 

\medskip\noindent
For the record I put other thoughts I went through in solving this exercise.

Recall from \cite[p. 146]{gri} that the normal bundle
$\mathcal{N}$ of a hypersurface of a smooth variety $X$ satisfies
$\mathcal{N}^\vee\cong\mathcal{O}_X(-C)|_{C}$. Taking duals we get that
$\mathcal{N}\cong\mathcal{O}_X(C)|_{C}$.

By adjunction formula $2g-2=(\mathcal{O}(C),\mathcal{O}(C))$. 
Applying Riemann-Roch
to $\mathcal{O}(C)$ (which is by definition the dual of the ideal sheaf of $C$,
which is a line bundle on $X$), we obtain that
$\chi(\mathcal{O}(C))=2+\frac{1}{2}(\mathcal{O}(C),\mathcal{O}(C))$. Then
$\chi(\mathcal{O}(C))=g+1$.

Now recall that $\chi(\mathcal{O}(C))=h^0(\mathcal{O}(C)
-h^1(\mathcal{O}(C))+h^2(\mathcal{O}(C))$. By Serre duality and $X$ being a K3
surface we see that $h^2(\mathcal{O}(C))\cong h^{0}(\mathcal{O}(-C))$, which is
the ideal sheaf of $C$. Any section of such a sheaf would vanish along $C$, and
since $X$ is compact we conclude there cannot be any such section.

Now we show that also $h^1(\mathcal{O}(C))=0$ to conclude that
$h^0(\mathcal{O}(C))=g+1$. 

We thus conclude that $h^0(\mathcal{O}(C))$




Since $C$ is smooth we can use the normal exact sequence
$$
\xymatrix{
0\ar[r]&\mathcal{T}_C\ar[r]&\mathcal{T}_X|_{C}\ar[r]&\mathcal{N}\ar[r]&0
}
$$
Taking Euler characteristic we see that
$\chi(\mathcal{T}_C)+\chi(\mathcal{N})=\chi(\mathcal{T}_X|_{C})$.


This means that we should be done once we compute $\chi(\mathcal{T}_X|_{C})$. 
For this we can use Riemman-Roch formula for coherent sheaves on a curve, which
tells us that
$$
\text{deg}(\mathcal{T}_X|_{C})=\chi(\mathcal{T}_X|_{C})
-\text{rk}(\mathcal{T}_X|_{C})\cdot\chi(\mathcal{O}_C)
$$
But of course we know that since $C$ is a curve, by Serre duality we get that
$\chi(\mathcal{O}_C)=1-g$. (Indeed:
$h^1(\mathcal{O}_C)=h^0(\Omega^1_C):=p_a(C)$.) And now the question is what is
the degree. Apparently this is just the first Chern class $c_1$ of the
restricted tangent bundle. So what is it? And what is $h^0(\mathcal{T}_C)$ if
the genus is 0 or 1, and that's it.

I
know that $h^0(\mathcal{T}_X)=0$ and $h^1(\mathcal{T}_X)=20$ by $X$ being a K3
surface, but I'm not sure what happens when we restrict to $C$.

If $g \geq 2$ we know that $h^0(\mathcal{T}_C)=0$, so that
$\boxed{\chi(\mathcal{T}_C)=h^1(\mathcal{T}_C)}$.
\bigskip
To compute the
latter Euler characteristic, which is given by definition by
$\chi(\mathcal{T}_X|_{C})=h^0(\mathcal{T}_X|_{C})-h^1(\mathcal{T}_X|_{C})$, we
first note that $h^0(\mathcal{T}_X|_{C})=0$, because this is the dimension of
the global holomorphic vector fields on $X$
restricted to $C$, which is constant along $C$ since $X$ is
smooth. And then this number is actually zero by Hodge numbers of a K3 surface.

and taking cohomology long exact sequence we obtain
$$
\xymatrix{
\cdots\ar[r]&H^{0}(T_X)\ar[r]&H^{0}(\mathcal{N})\ar[r]&H^{1}(T_C)\ar[r]&\\
H^{1}(T_X)\ar[r]&H^{1}(\mathcal{N})\ar[r]& \cdots
}
$$
\medskip\noindent
Or is it $h^1(\mathcal{N})$? See \cite{HarrMorr}. If this was the case, then we
can use adjunction formula as above to get that
$2g-2=(\mathcal{N},\mathcal{N})=\text{deg}_C \mathcal{N}$. Then we may find
$h^0(\mathcal{N})$ via Riemann-Roch:
$$
h^0(\mathcal{N})-h^0(K_C-\mathcal{N})=\text{deg}\mathcal{N}-g+1
$$
Note that $h^0(N_C-\mathcal{N})=h^1(-K_C+N+K_C)=h^1(\mathcal{N})$ via Serre
duality, and by Riemann-Roch on a surface as above we see that
$$
g+1=\chi(\mathcal{N})=h^0(\mathcal{N})-h^1(\mathcal{N})
\implies h^1(\mathcal{N})=-g-1+h^0(\mathcal{N})
$$
so that
\begin{align*}
h^0(\mathcal{N})-(-g-1+h^0(\mathcal{N}))&=\text{deg}\mathcal{N}-g+1\\
\implies \text{oops! I lost $h^0(\mathcal{N})$ in this operation…}
\end{align*}
Maybe if I just use normal exact sequence and realise that
$$
h^0(\mathcal{N})=h^0(\mathcal{T}_X|_{C})-h^0(\mathcal{T}_C)
$$
I know that $h^0(\mathcal{T}_C)=0$ for $g>1$, so the question is how to compute
the restricted holomorphic vector fields.
%\end{enumerate}
\end{proof}

\section{Deformations of Stanley-Reisner schemes}
\label{section-sr}

\noindent
My aim in this section is to cover the minimal
theory needed to apply results in \cite{jan2}
to my deformation problem. The main theorem
I expect to use is the following:

\begin{theorem}
\label{theorem-deformations-threefolds}
\begin{reference}
\cite[Theorem 5.7]{jan2}
\end{reference}
If $\mathcal{K}$ is a 3-dimensional manifold then
$$
\dim T^1_{\mathbb{P}(K)}
=11d_3+5e_3+3e_4+e_{\geq 5}+c_{\geq 6}+5f_1^{(3)}+2f_1^{(4)}+h^2(\mathcal{K}).
$$
\end{theorem}

\noindent
I need to do some interpretation first.
From \cite[p. 1]{jan3}:

\begin{quotation}
  ``For each $\mathbb{C}$-algebra $A$,
there is a cohomology theory providing modules $T^i_A$.
However, only $T^1_A$ and $T^2_A$ are relevant for the deformation
theory of $\Spec A$. The module $T^1_A$ collects the infinitesimal
deformations of $A$ and $T^2_A$ contains the obstructions
for lifting these deformations to decent parameter spaces.
Eventually, $\text{Der}_{\mathbb{C}}(A,A)$ will be called $T_A^0$."
\end{quotation}

\cite[p. 3]{jan3}:
\begin{quotation}
``In general, for a finitely generated $\mathbb{C}$-algebra $A$,
the modules $T_A^i$ allow the following ad hoc definitions:
Let $P=\mathbb{C}[x_0,…,x_n]$ mapping onto $A$ so that $A \simeq P/I$
for an ideal $I$. The $T_A^1$ is the cokernel of the natural map
$\text{Der}_{\mathbb{C}}(P,P) \to \Hom_P(I,A)$.
Moreovr, […] and we obtain $T^2_A$ as the cokernel
of the induced map $\Hom_P(P^m,A)\to\Hom_A(R/R_0,A)$.''
\end{quotation}

\begin{remark}[dani]
\label{remark-dani1}
Let's think more about that definition of $T^1_A$
as the cokernel of the ``natural map''
$\text{Der}_\mathbb{C}(P,P) \to \Hom_P(I,A)$.

Pick a derivation $\chi \in \text{Der}_{\mathbb{C}}(P,P)$.
We want a map $I \to A=P/I$.
Pick an element $a \in I$ and map it to $\chi(a)$.
And that's it. It might be zero or not.
If the $\chi$ we picked maps any $a \in I$ to some element in $I$,
then this $\chi$ is in the kernel of the natural map.
So $T^1_A$ is the image of the natural map modulo the kernel.
\end{remark}

\noindent
Unfortunately this doesn't say much about the following crucial fact:
the $T^i_A$ modules are $\mathbb{Z}^{n+1}$-graded.
But here goes an explanation.
We use the notation that $e_p$ is a ``face not in the complex'',
which of course corresponds to a generating monomial $x^p$
of the SR ideal $I$.
It has a degree, of course (given by the ``support'' of $p$).
Now here's a rather sloppy step but:
since the cotangent modules are built upon $I$,
who is given by $e_p$,
we just notice that the degree of $e_p$
along with a (very natural, too) notion of degree among relations, 
induce a grading in all the cotangent modules $T_A^i$.
Thus, the grading is morally just the degree of the monomials,
which is in fact the support of $p$, a vector $\mathbf{c} \in \mathbb{Z}^{n+1}$.

Next in line is the decomposition $\mathbf{c}=\mathbf{a}-\mathbf{b}$
into positive minus negative part. But how exactly…?

\section{Miscellaneous}
\label{section-misc}

Look for the ``Perfect obstruction theory''.
This is used by Bruzzo for deformations of (super)stacks,
work with E. Paiva and ---.


\end{document}
