\input{preamble}

\begin{document}

\title{Representation Theory}
\maketitle

\phantomsection
\label{section-phantom}
\hfill
\href{http://github.com/danimalabares/stack}{github.com/danimalabares/stack}

\tableofcontents

\section{Basic definitions}
\label{section-basic-definitions}

\begin{definition}
\label{definition-representation}
A {\it representation of Lie algebra} $\mathfrak{g}$ is a vector space $V$ 
and homomorphism
$$
\rho:\mathfrak{g} \to \text{End}(V)
$$
of Lie algebras, i.e.,
$$
\rho([x,y])=\rho(x)\rho(y)-\rho(y)\rho(x).
$$
\end{definition}

\begin{remark}
\label{remark-representations-are-modules}
Another name for a representation of a Lie algebra $\mathfrak{g}$ is a 
{\it $\mathfrak{g}$-module}. This is like the abstract definition of an algebra
using a morphism. It's as if the morphism lets us define a sort of product of
elements in the codomain by elements in the domain.
\end{remark}

It is not possible to classify all representations of a Lie algebra
$\mathfrak{g}$. But there is a theorem by Weyl that says that if $\mathfrak{g}$
is finite-dimensional and (semi)simple, then every finite-dimensional
representation of $\mathfrak{g}$ 
(i.e. a representation where $V$ is finite-dimensional) 
is isomorphic to a direct sum of irreducible representations 


\section{Representations of SL(2,C)}
\label{section-sl2c}

\noindent
Klein classified nontrivial finite subgroups of $\text{SL}_2(\mathbb{C})$ 
up to conjugacy:
\begin{itemize}
\item Cyclic groups $\mathcal{C}_m$, $m \geq 2$. $|\mathcal{C}_m|=m$,
$$
\mathcal{C}_m=\left\{\begin{pmatrix}
\zeta&0\\ 
0&\zeta^{-1}\end{pmatrix}:\zeta^m=1
\right\}.
$$
\item Dihedral groups $\mathcal{D}_m$, $m\geq 2$,
$$
\mathcal{D}_m=\left<\mathcal{C}_{2m},\begin{pmatrix}
0&-1\\ 
1&0
\end{pmatrix}\right>.
$$
For example, $\mathcal{D}_2$ is the {\it quaternion group} 
$\left<\underbrace{\begin{pmatrix}
i&0\\ 
0&-i
\end{pmatrix}}_{i},\underbrace{\begin{pmatrix}
0&-1\\ 
1&0
\end{pmatrix}}_{j}\right>$.

\item Binary tetrahedral group $\mathcal{T}$, with $|\mathcal{T}|=24$.
\item Binary octahedral group, $\mathcal{O}$, with $|\mathcal{O}|=48$.
\item Binary icosahedral group $\mathcal{I}$ with $|I|=120$.

These three are related to the rotation groups of the
Platonic solids by the 2-to-1 covering
$\text{SU}_2 \to \text{SO}_3(\mathbb{R})$.
To classify the finite subgroups of $\text{SL}_2(\mathbb{C})$:
first show that any such subgroup is conjugate to a subgroup
of $\text{SU}_2$ so that its image is a finite subgroup of
$\text{SO}_3(\mathbb{R})$. We can classify those by considering their
action on the sphere and showing that they have to be the
rotation group of one of the Platonic solids or a dihedral group
or a cyclic group (related to planar polygons).
\end{itemize}

\noindent
At some point of the XX century it was realised that this
classification is an instance of an ADE classification.

Now we explain the (historically second explanation of this) McKay
correspondence.
If $\Gamma$ is a nontrivial finite subgroup of $\text{SL}_2(\mathbb{C})$,
define the  {\it McKay graph} $\tilde{\Delta}_\Gamma$ as follows:
the vertices are $0,1,2,\ldots,\ell$ which index the
irreducible representations $R_0,R_1,\ldots,R_\ell$ of $\Gamma$.

Let $Q$ be the 2-dimensional $\Gamma$ coming from
$\Gamma<\text{SL}_2(\mathbb{C})$.
(Note that $Q$ is one of the $R_i$'s unless $\Gamma=\mathcal{C}_m$,
in which case $Q$ is reducible.)
The number of edges between $i$ and $j$ in $\tilde{\Delta}_\Gamma$ 
is the multiplicity of $R_j$ in $Q \otimes R_i$.
So: take the representation $R_i$, tensor it with the
2-dimensional representation $Q$, 
express it as a direct sum of irreducible representations
and perhaps $R_j$ will appear with nonzero multiplicity,
in which case you put $R_j$ with that number of edges.
So in general it's a multigraph, though in most
cases its a simple graph.
Also note that this multiplicity coincides
with the multiplicity of $R_i$ in $Q \otimes R_j$
since $Q^* \cong Q$; that is the graph is not directed.

The surprising thing is that when we do this we obtain
the affine Dynkin diagrams. McKay's famous observation says:
\begin{enumerate}
\item $\tilde{\Delta}_\Gamma$ is a simply-laced affine Dynkin
diagram.

\item This is a bijection:
$$
\left\{\substack{\text{nontrvial finite} \\ \Gamma \subset
\text{SL}_2(\mathbb{C})}\right\}\Big/\text{conj.}
\xymatrix{\ar@{<~>}[r]&}
\left\{\substack{\text{simply-laced} \\ \text{affine Dynkin diagrams}}\right\}.
$$
\end{enumerate}

\noindent
The correspondance reads
\begin{align*}
\mathcal{C}_m &\xymatrix{\ar@{<->}[r]&}
\widetilde{A}_{m-1}\\
\mathcal{D}_m &\xymatrix{\ar@{<->}[r]&}
\widetilde{D}_{m+2}\\
\mathcal{T}&\xymatrix{\ar@{<->}[r]&}\tilde{E}_6\\
\mathcal{O}&\xymatrix{\ar@{<->}[r]&}\tilde{E}_7\\
\mathcal{I}&\xymatrix{\ar@{<->}[r]&}\tilde{E}_8.
\end{align*}

\medskip\noindent
Now we explain the geometric approach.
Given a finite $\Gamma \subset \text{SL}_2(\mathbb{C})$, consider
the quotient $Y_\Gamma=\Gamma\setminus Q$ as an algebraic variety,
that is, (since $Q$ is just a representation of $\Gamma$, i.e.
a 2-dimensional affine space with an action of $\Gamma$)
we can take its quotient in the sense described before:
$Y_\Gamma$ is the affine variety with coordinate algebra
$\mathbb{C}[x,y]^\Gamma$.

\begin{lemma}
\label{lemma-action-is-free-outside-origin}
The action of $\Gamma$ is free on $Q\setminus\{0\}$ is free.
\end{lemma}



\section{Borel-Weil-Bott theorem}
\label{section-Borel-Weil-Bott-thoerem}

[These are {\it very} sketchy notes after a conversation with Lyalya Guseva on
Borel-Weil-Bott theorem]

Borel-Weil-Bott theorem is a device for computing cohomologies of sheaves.

Take the Grassmanian $\text{Gr}(n,k)$, which may be obtained as a quotient $G$
of a group $G$ by a parabolic subgroup $P$. We would like it if $P$ was a
semisimple group since those are classified, but unfortunately it is not. 
So instead we use so-called Levi quotient denoted by $L$ which is semisimple
and allows us to understand the cohomologies of the variety $G/P$.

Fortunately the representation theory of $\text{Gr}(k,n)$ is well known, in fact
we get $\text{GL}_k \times \text{GL}_{n-k}$.

There are functors associated to $L$, which are described by a sequence of
numbers $a_1,\ldots,a_n$. Along with other sequences of numbers
$k_1,\ldots,k_\ell$, and $\rho$ (the latter is a concept in representation
theory but we may ultimately think of it as another sequence of numbers) we may
construct an action of the symmetric group $S^n$ acting on these sequences and
obtain a result concerning the cohomology $H^{\ell(\sigma)}(L)$, and in
particular we find that if two entries in our list of numbers coincide, the
cohomology will vanish.


\bibliography{my}
\bibliographystyle{amsalpha}

\end{document}

