\input{preamble}
\begin{document}

\title{A tour through algebraic surfaces}
\maketitle

\noindent
Minicourse by 
Aline Zanardini
(EPFL, Switzerland),
IMPA Summer School, 2026.

\medskip\noindent

\hfill Notes at 
\href{http://github.com/danimalabares/stack}
{github.com/danimalabares/stack}

\bigskip\noindent

\tableofcontents

\section{Plan}
\label{section-plan}

\begin{enumerate}
\item Basics.
\item Birational maps and classification.
\item Elliptic surfaces.
\item Halphen surfaces.
\item K3 surfaces and friends.
\item Research questions.
\end{enumerate}

\section{References}
\label{section-references}

\begin{enumerate}
\item A. Beauville. Complex algebraic surfaces.
\item R. Miranda. Overview of algebraic surfaces.
\item M. Reid. Chapters on algebraic surfaces.
\end{enumerate}

\section{Introduction}
\label{section-intro}

\noindent
Everything will be over the complex numbers.
By ``surface'' we mean ``algebraic surface'',
but we drop the term ``algebraic''.
Informally, a surface is a smooth projective algebraic variety of dimension 2.
We may think of surfaces also as
compact connect complex manifolds of dimension 2.

$\mathcal{M}(S)$, meromorphic functions on $S$, has transcendence degree
2 over $\mathbb{C}$.
That is, for $p,q \in S$
there exists $f \in \mathcal{M}(S)$
such that $f(p)\neq f(q)$
and for all $p \in S$ there exists $f,g \in \mathcal{M}(S)$
such that $(f,g)$ give local coordinates at $p$.

\begin{remark}
\label{remark-transendence-implies-surface-def}
The remark on transcendence just made implies that $S$
is smooth algebraic of dimension 2.
Indeed, compactness implies that there exists
meromorphic functions  $\varphi_1,…,\varphi_n \in \mathcal{M}(S)$
such that $S \xymatrix{\ar@{^{(}->}[r]^{alg}&}\mathbb{P}^n$,
$p \mapsto (1:\varphi_1(p):…:\varphi_n(p))$ and Chow theorem
imply that $S = V(f_1,…,f_k)$ with 
$f_i \in \mathbb{C}[x_0,x_1,…,x_n]$.
\end{remark}

\begin{example}
\label{example-surfaces}
\begin{enumerate}
\item $S = V(f) \subset \mathbb{P}^3$
with $f \in \mathbb{C}[x_0,x_1,x_2,x_3]$
homogeneous of degree $d$.
For example, $f=x_0^d+x_1^d+x_2^d+x_3^d$.
Take $d=1$, then we can assume that $f=x_i$,
and then $S\simeq\mathbb{P}^2$.
If $d=2$ we can assume that  $f=x_0x_1-x_2x_3$ and
one can verify that $S \simeq \mathbb{P}^1 \times \mathbb{P}^1$.
If $d=3$, $S \simeq Bl_6\mathbb{P}^2$.
These are all rational surfaces.

If $d=4$, $S$ is a K3 surface.

\item (Complete intersections of type $(d_1,…,d_{n-1})$ in $\mathbb{P}^n$.)
Let $f_1,…,f_{n-2} \in \mathbb{C}[x_0,x_1,…,x_n]$ homogeneous
of degrees $d_1,…,d_{n-2}$.
Then $S=V(f_1,…,f_{n-2})$.

E.g., intersections of type $(2,3)$ in  $\mathbb{P}^4$
or of type $(2,2,2)$ in $\mathbb{P}^5$, which are K3 surfaces.

\item (Ruled surfaces.)
Take any curve $C$ and consider  $C \times \mathbb{P}^1$
(or anything birational to it).
[This is interesting because we can consider the projection to $C$,
and we get a fibration.]
\end{enumerate}
\end{example}

\section{Curves on surfaces (divisors)}
\label{section-curves-on-surfaces}

\noindent
We will use: $\text{Div}(S)/\sim \simeq \text{Pic}(X)=H^1(S,\mathcal{O}_S^*)$
where $\sim$ is linear equivalence.

$D \subset S$ is a divisor if and only if $D$ is one of the following:
\begin{enumerate}
\item (Cartier divisor.)
[I can think of $D$ given locally as the zero locus of some function.]
$D:f=0$ for  $0 \neq  f \in \mathcal{M}(S)$ + gluing condition.
This is equivalent to saying that $D|_f=\text{div}(f)$.

\item (Weil divisors.)
$D=\sum n_iC_i$ for $n_i \in \mathbb{Z}$,
$C_i \subset S$ irreducible curves,
where there are only finitely many nonzero coefficients $n_i$.

\item (Line bundle.)
$D$ effective (i.e. $n_i\geq 0$), we associate
$\mathcal{O}_S(D)$, a holomorphic line bundle on $S$ along
with a nonzero section [whose zero locus defines $D$.
Recall that $U \mapsto  \mathcal{O}_S(D)(U)$ is ``given''
by $g(z,w)$ such that  $fg$ is holomorphic
where $D|_U=\text{div}(f)$.
And we extend by linearity.
\end{enumerate} 

\noindent
[And we say that two divisors are linearly
equivalent if their difference is $div(f)$ for some $f$.]

\begin{example}
\label{example-divisors}
\begin{enumerate}
\item $\text{Pic}(\mathbb{P}^2)=\mathbb{Z}$
generated by $\mathcal{O}_{\mathbb{P}^2}(1)$,
the dual of the tautological bundle, or, equivalently,
the class of a hyperplane $H$,
whose sections are polynomials of degree 1
(in the appropriate number of variables).

\item $\text{Pic}(\mathbb{P}^n)=\mathbb{Z}$ 
is generated by $\mathcal{O}_{\mathbb{P}^n}(1)$.
We have $\mathcal{O}_{\mathbb{P}^n}(d):=\mathcal{O}_{\mathbb{P}^n}^{\otimes d}$.
For $n=2$, the sections of $\mathcal{O}(d)$ are
plane curves of degree 2.
\end{enumerate}
\end{example}

\noindent
Suppose that $\pi:S \xymatrix{\ar@{^{(}->}[r]&} \mathbb{P}^n$.
The {\it hyperplane class} is  $H|_S:= \pi^*\mathcal{O}_{\mathbb{P}^n}(1)$.
We denote $\mathcal{O}_S(1)=\mathcal{O}_S(H|_S)$.
This is always a non-trivial class.

We also have the {\it canonical class},
which is the class of any divisor $K$ 
such $\mathcal{O}_S(K)=\Omega^2_S$,
the sheaf of holomorphic 2-forms,
informally its elements look like $f(z,w)dx\wedge dw$.
We denote this as $K$,  $K_S$ or $\omega_S$.

\begin{example}[Canonical class of projective plane]
\label{example-canonical-class}
Take coordinates $Z=X_1/X_0$ and $W=X_2/X_0$
on $U_0: X_0 \neq 0$.
Likewise put $U=X_0/X_1$, $V=X_2/X_1$ at $U_1$.
Then $dZ\wedge dW=-U^{-3}dU\wedge dV$.
Thus, the canonical class of $\mathbb{P}^2$
is linearly equivalent to $-3H$.
In general, $K_{\mathbb{P}^n}\sim -(n+1)H$.

In practice, we can use this to compute $K_S$
for any  $S \xymatrix{\ar@{^{(}->}[r]&} \mathbb{P}^n$
via $\omega_S=\omega_{\mathbb{P}^n}\otimes \Lambda^{n-2}N_S/\mathbb{P}^n$.
\end{example}


\section{Numerical invariants}
\label{section-numerical-invariants}

\begin{itemize}
\item (Hodge decomposition.)
$$
H^k(S,\mathbb{C})=\bigoplus_{p+q=k}\underbrace{H^q(S,\Omega^p_S)}_{
H^{p,q}(S)}.
$$
Setting $h^{p,q}=\dim H^{p,q}$ + symmetries,
we have the Hodge diamond
$$
\begin{aligned}\xymatrix@C=.5em@R=.5em{
&  &   h^{0,0}\\
&  h^{0,1}&  &  h^{1,0}\\
h^{0,2}&&  h^{1,1}&  &  h^{2,0}\\
&  h^{1,2}&  &  h^{2,1}\\
&  &  h^{2,2}
}\end{aligned}
\qquad =\qquad 
\begin{aligned}\xymatrix@C=.5em@R=.5em{&  &   1\\
&  q&  &  q\\
p_g&&  h^{1,1}&  &  p_g\\
&  q&  &  q\\
&  &  1
}\end{aligned}
$$
where we call $q$ the {\it irregularity}
$p_g$ the {\it geometric genus}
and the {\it Euler characteristic}
is the alternating sum of Betti numbers.
Notice that $h^{1,1}\geq 1$.
\end{itemize}

\section{The Néron-Severi group and Lefschetz (1,1) theorem}
\label{section-neron-severi}

\noindent
There is always a map
\begin{align*}
c_1: \text{Pic}(S) &\longrightarrow H^2(S,\mathbb{Z})
\overset{\substack{\text{Poincaré} \\ \text{dual}}}{\cong}
H_2(S,\mathbb{Z}) \\
\mathcal{L} &\longmapsto c_1(\mathcal{L})
\end{align*}

\noindent
We define $\text{NS}(X):=\text{Im}c_1/\text{torsion}$.
It is a finitely generated abelian group isomorphic to $\mathbb{Z}^{\rho}$
where $\rho$ is the Picard rank.

%\begin{theorem}[Lefschetz (1,1)-classes]
%\label{theorem-lefschetz-11}

\begin{theorem}
\label{theorem-form}
$$
\text{NS}(S)=H^2(S,\mathbb{Z})\cap H^{1,1}(S)
$$
$\text{NS}(S)$, $\text{Div}(S)/\sim$ and $\text{Pic}(S)$
have more structure!
There exists a unique bilinear form 
\begin{align*}
\cdot: \text{Div}(S)\times \text{Div}(S) &\longrightarrow \mathbb{Z} \\
(C,D) &\longmapsto C\cdot D
\end{align*}

\noindent
such that 
\begin{enumerate}
\item If $C $ and $D$ are smooth curves which intersect
transversally, then $C.D=\# C \cap D$.
\item If $C \sim C'$, then $C\cdot D = C' \sim D$.
\end{enumerate}
\end{theorem}

\begin{definition}
\label{definition-form}
$C.D=\mathcal{O}_S(C).\mathcal{O}_S(D)
=\left<c_1(\mathcal{O}_S(D) \cup c_1(\mathcal{O}_S(D)),
[S]_{\text{fund}}\right>$,
and we say that $C.C=\text{deg} \mathcal{N}_{C/S}$.
\end{definition}

\noindent
[An important invariant of a surface
is the self-intersection number of the canonical class.]
Particular case: $K^2=c_1^2$
where $c_1^2$ is the first Chern number.

\begin{example}
\label{example-intersection}
\begin{enumerate}
\item $K_{\mathbb{P}^2}^2=(-3 H).(-3 H)=9$.
\item (Degree of $S$.)
Let $S \xymatrix{\ar@{^{(}->}[r]&} \mathbb{P}^n$.
Recall our notation above for
$\mathcal{O}_S(1).\mathcal{O}_S(1)$.
If $n=3$,
the degree $d$ of $S$ is the degree of $\mathcal{N}_{S/\mathbb{P}^3}$.
Thus the adjunction becomes $\omega_S=\mathcal{O}_S(d-4$

\item $C$ smooth curve, then 
$f:S \xymatrix{\ar@{->>}[r]&}C$,
$F$ fiber, then  $F^2=0$.

\item $g=\tilde{S}\overset{d}{\to}D$,
then $D_1,D_2 \in \text{Div}(S)\implies 
g^*(D_1)\cdot g^*(D_2)=d(D_1.D_2)$.
\end{enumerate}
\end{example}

\begin{definition}
\label{definition-nef-and-ample}
A divisor $D \subset S$ is {\it nef} if and only if
for each $C \subset S$ irreducible,
$D\cdot C \geq 0$.
We say that $D$ is ample if and only if
$D^2>0$ and for each irreducible $C \subset S$
we have $D.C \geq 0$.
\end{definition}

\section{Big theorems}
\label{section-big-theorems}

\begin{theorem}
\label{theorem-}
Let $S$ be a surface.
\begin{enumerate}
\item (Noether's formula.)
$$
12\chi = K^2+e=c_1^2+c_2
$$
where $e$ is the Euler number.

\item (Riemann-Roch.) Given $D \in \text{Div}(S)$,
$$
\chi(\mathcal{O}_S(D))=\frac{D.(D-K)}{2}+\chi(\mathcal{O}_S).
$$

\item (Genus formula.) Given $C \subset S$ smooth,
$$
2g-2=C.(C+K).
$$
\end{enumerate}
\end{theorem}

\begin{example}
\label{example-big-theorems}
\begin{enumerate}
\item For $\mathbb{P}^2$, $K^2=9$ and $e=3$.
Thus $12\chi = 12 \implies \chi =1$.
The irregularity is $q=H^0(\mathbb{P}^2,\Omega^1)=0$.
The arithmetic genus is $p_g=\dim H^0(\mathbb{P}^2,\Omega^2)=0$.
Thus, the Hodge diamond is
$$
\xymatrix@C=.5em@R=.5em{
& & 1\\
& 0 && 0\\
0 && 1 && 0\\
& 0 && 0\\
& & 1
}
$$
\item Any rational surface has a Hodge diamond
$$
\xymatrix@C=.5em@R=.5em{
& & 1\\
& 0 && 0\\
0 && \rho && 0\\
& 0 && 0\\
& & 1
}
$$
The Noether formula gives
$12=K_S^2+\rho+2 \implies K_S^2=10-\rho$.

\item $C \subset \mathbb{P}^2$ smooth of degree $d$,
then $2g-2=d^2-3d$ and
$g=\frac{d^2-3d+2}{2}=\frac{(d-1)(d-2)}{2}
=\binom{d-1}{2}$.
In particular,
if $d=1$ or  $2$,
then  $g=g(C)=0$ 
if $d=3$ then  $g(C)=1$.
\end{enumerate}
\end{example}

\section{Birational maps and classification theorem}
\label{section-birational-maps-and-classification-theorem}

\noindent
Question: given two surfaces $S$ and $S'$,
when are $S$ and $S'$ birational/biregular or isomorphic?

\begin{itemize}
\item Rules surfaces
(all birational to $C\times \mathbb{P}^1$ for some $C$.
\item Classify minimal (models of) surfaces
by looking at the positive of $K$ 
(and related invariants
$q$, $p_g$ and $\kappa$).
\end{itemize}

\begin{center}\begin{tabular}{c|c|c|c|c}
$K$ & $\kappa$ & $p_g$ &  $q$  & Structure\\
\hline\hline
$K^2>0$&2&&&General type\\ \hline
$K^2=0$&1&&&Elliptic*\\\hline
$K=0$&0&
$\substack{1\\0\\1\\0 \\ }$&
$\substack{2 \\1\\0\\0 }$
&$\substack{
\text{abelian} \\ 
\text{hyperelliptic}\\
\text{K3}\\
\text{Enriques}}$\\\hline
&$-\infty$&$\substack{0 \\ 0}$&
$\substack{\geq 1 \\ 0}$&
$\substack{\text{ruled surfaces} \\ \text{rational}}$
\end{tabular}\end{center}

\medskip\noindent
The following proposition is also a definition:

\begin{proposition}
\label{proposition-birational}
Let $S$ and $S'$ be two surfaces, then the following
statements about $S$ and $S'$ are equivalent:
\begin{enumerate}
\item $S$ and $S'$ are birational
\item $\mathcal{M}(S)\simeq \mathcal{M}(S')$
as algebras (over $\mathbb{C}$).
\item There exists $U \subset S$, $V \subset S'$ 
opens such that $U \simeq V$.
\end{enumerate}
\end{proposition}

\section{The blow up of the projective plane at a point}
\label{section-blow-up}

\noindent
Consider $p=(0:0:1)$ and
\begin{align*}
\varphi_p: \mathbb{P}^2\setminus \{p\} &\longrightarrow \mathbb{P}^1 \\
(x_0:x_1:x_2) &\longmapsto (x_0:x_1)\\
q &  \longmapsto L_{pq}
\end{align*}

\noindent
Consider also
$$
\Gamma:=\left\{((x_0:x_1:x_2),(y_0:y_1)) \in \mathbb{P}^2\times \mathbb{P}^1
: \det\begin{pmatrix}x_0&x_1\\ y_0&y_1\end{pmatrix}=0\right\}.
$$

By construction,
$$
\pi:\Gamma \to \mathbb{P}^2
$$
is an isomorphism over $U=\mathbb{P}^2 \setminus \{p\}$
and $\pi^{-1}(p)=\{p\}\times \mathbb{P}^1 \simeq\mathbb{P}^1:=E$.

\begin{definition}
\label{definition-blow-up}
$\pi:\Gamma \to \mathbb{P}^2$ as above is called
the {\it blowup} of $\mathbb{P}^2$ at $p$ 
(also $\Gamma$). The curve $E$ is called
the {\it exceptional divisor}.
\end{definition}

\noindent
Note that we have the following local description
(chart $x_2=1$):
$\{(x,\ell) \in \mathbb{C}^2\times \mathbb{P}^1: x \in \ell\}
\to \mathbb{C}^2$.
That is, the exceptional divisor may be identified with
the zero section of the tautological bundle $\mathcal{O}_{\mathbb{P}^1}(-1)$,
which means that the self intersection of $E$ is 1,
i.e. $E^2=1$.

\begin{remark}
\label{remark-blowup}
\begin{enumerate}
\item The surface $Bl_p\mathbb{P}^2:=\Gamma$
is the Hirzebruch surface $\mathbb{F}_1$ (or $\Sigma_1$).

\item Choosing local coordinates, we can talk about 
$Bl_p S$ for $p \in S$.

\item In fact, $\pi:Bl_pS \to S$ 
is characterized by the following universal property:
if $f:\tilde{S}\to S$ is birational such that $f^{-1}$
is not defined at $p \in S$, then $f$ factorizes
as $f:\tilde{S} \underbrace{\xrightarrow{g}}_{\text{bir}} 
\hat{S}:=Bl_pS \xrightarrow{\varepsilon} S$ 
where $\varepsilon^{-1}(S\setminus \{p\}) \simeq S \setminus \{p\}$ 
and $\varepsilon^{-1}(p) \simeq \mathbb{P}^1$.
\end{enumerate}
\end{remark}

\begin{proposition}
\label{proposition-blowup}
Consider $\pi:\tilde{S}=Bl_pS \to S$
as above, with exceptional divisor $E$,
\begin{enumerate}
\item $E^2=-1$,

\item $E\cdot \pi^*D=0$ for all $D \in \text{Div}(S)$,

\item If $C \subset S$ has mutliplicity $m$ at $p$, then
$\tilde{C}:= \overline{\pi^{-1}(C) \setminus E}\subset\tilde{S}$
satisfies
$\tilde{C}\cdot E=m$ and $\tilde{C}^2=C^2-m^2$

\item $K_{\tilde{S}}=\pi^* K_S+E\implies K_{\tilde{S}}^2=K_S^2-1$ 

\item $q,p_g$ and  $\chi$ are the same for $\tilde{S}$ as for $S$ 

\item $e$,  $h^{1,1}$ and $b_2$ all go up by 1

\item $\text{Pic}(\tilde{S})=\pi^*(\text{Pic}(S))\oplus \mathbb{Z}E$.
\end{enumerate}
\end{proposition}

\section{Castelnuovo's criterion}
\label{section-castelnuovo-criterion}

\noindent
[It turns out that if you find a surface with a $-1$ curve,
then that surface is the blow up of another!]

\begin{theorem}
\label{theorem-castelnuovo}
If on a surface $\tilde{S}$ one finds a curve $E \simeq \mathbb{P}^1$
with $E^2=-1$, then $\tilde{S}=Bl_pS$
for some $p \in S$.
\end{theorem}

\begin{definition}
\label{definition-minimal-surface}
A surface $S$ is {\it minimal} 
if it has no $(-1)$-curves
(i.e. a curve $\simeq \mathbb{P}^1$ with self-intersection -1).

(If and only if
$f:S \xymatrix{\ar@{-->}[r]&}\tilde{S}$ and $S$ minimal,
then $f$ is an isomorphism.)
\end{definition}


\section{Mori's point of view}
\label{section-moris-point-of-view}

\noindent
If $S$ is a surface such that $K_S$ is not nef
[so, by definition of nef there must exist
an irreducible curve that intersects
the canonical divisor negatively, and this curve
in fact is a $(-1)$ curve:]
then there exists a $(-1)$-curve on $S$.




\section{Kodaira dimension}
\label{section-kodaira-dimension}

\noindent
Let $S$ be a surface.

\begin{definition}
\label{definition-kodaira-dimension}
For each $n \geq 0$,
we define the {\it pluri-genera}
$$
P_n:= \dim H^0(S,\mathcal{O}_S(n K))
$$
and the {\it Kodaira dimension}
$$
\kappa:=
\begin{cases}
-\infty&\qquad \text{if }P_n=0\forall n \\
\text{smallest $k$ such that }
P_n=O(n^k)&\qquad \text{otherwise}
\end{cases}
$$
\end{definition}

\begin{remark} \label{remark-kodaira} (By Daniel.) In one of Misha's course we
have that the function $\dim H^0(n K)$ is known to be polynomial for all
projective varieties (Birkar, Cascini, Hacon, McKernan). Conjecturally it is
always polynomial. Thus, the definition of Kodaira dimension is nothing
more than the degree of this polynomial.

Further, in \cite[Definition 2.2.26]{huc} we see
that this should be equivalent to defining it as the
transcendence degree over $\mathbb{C}$
of the fraction field of the ring 
$\bigoplus_{m \geq 0}H^0(X,K_X^{\oplus m})$,
which is endowed with a natural product which, in general for
vector bundles $E$ and $F$, maps
$H^0(X,E) \otimes H^0(X,F) \to H^0(X,E \otimes F)$.
\end{remark}

\section{Results of Mori's theory}
\label{section-results-of-mori}

\begin{theorem}
\label{theorem-one}
Given a surface $S$,
there exists a chain
$$
S=S^0 \xrightarrow{\sigma_1} S^1 \to \cdots \xrightarrow{\sigma_N}S_N=S'
$$
where each $\sigma_i$ is the contraction of a $(-1)$ curve $E_i$
and $S'$ satisfies
\begin{enumerate}
\item $K_{S'}$ is nef or
\item $S'$ is a $\mathbb{P}^1$-bundle
over a curve or 
\item $S' \simeq \mathbb{P}^2$.
\end{enumerate}
\end{theorem}

\begin{theorem}
\label{theorem-two}
If a surface $S$ is such that
$K_S$ is not nef then there exists $\varphi:S \to X$,
with $\dim X=0,1$ or $2$, contracting at least
one curve to a point, such that 
$-K_S\cdot C>0$ for every curve $C$ contained in a fiber of $\varphi$.
\end{theorem}

Note:
$$
\dim X=
\begin{cases}
0\xymatrix{\ar@{~>}[r]&}&-K \text{ is ample}\\
1\xymatrix{\ar@{~>}[r]&}&S \text{ is a conic bundle}\\
2\xymatrix{\ar@{~>}[r]&}& S=Bl_{p_1,…,p_n}.
\end{cases}
$$


\section{Examples and the cubic surface revisited}
\label{section-examples-and-cubic}

\begin{enumerate}
\item Slogan: ``blowups resolve singularities''.
Consider the plane cubic
$$
C=\{(x_0:x_1:x_2) \in \mathbb{P}^2:
\underbrace{x_1^2x_2-x_0^2(x_0+x_2)=}_{:=f_C}0\},
$$
the point $p=(0:0:1)$
and $\pi:Bl_p\mathbb{P}^2 \to \mathbb{P}^2$,
then
$$
\pi^{-1}(C)=
\begin{cases}
f_C=0\qquad & \\
x_0y_0=y_0x_1\qquad &
\end{cases}
$$
[This curve has a simple node.
In the blowup, the curve will intersect
the exceptional divisor in two points ---
we have ``separated'' the singularity into two points]

\item $Bl_1\mathbb{P}^1 \times \mathbb{P}^1
\simeq Bl_1 \mathbb{F}_1
\simeq Bl_2\mathbb{P}^2$.
[Consider two curves $L_1$ and $L_2$ on $\mathbb{P}^1 \times \mathbb{P}^1$.
Their strict transforms along with the exceptional divisor
become $(-1)$ curves.
Then contract $\tilde{L}_1$, this gives $\mathbb{F}_1$,
the Hirzebruch surface, which is just the blowup
at a point of $\mathbb{P}^2$.
$\mu(\tilde{L}_2)=-1$, $\mu(E\setminus \tilde{L}_1)=0$.]

\item Consider two cubics $C_1$ and $C_2$.
They intersect in 9 points.
Do
\begin{align*}
\mathbb{P}^2 &\xymatrix{\ar@{-->}[r]&} \mathbb{P}^1 \\
p &\longmapsto (C_1(p):C_2(p))
\end{align*}

\noindent
Let $S=Bl_{p_1,…,p_9}\mathbb{P}^2$.
Then
\begin{enumerate}
\item $K_S^2=0$ [Because $K^2$ of  $\mathbb{P}^2$ is 9,
and for every blowup we substract 1!]
\item $\text{Pic}(S)$  ``$=$'' $\pi^*\text{Pic}(\mathbb{P}^2)
\oplus\mathbb{Z}E_1 \oplus … \oplus \mathbb{Z}E_q$.
\item $-K_S=-\pi^* (K_{\mathbb{P}^2})
-E_1-…-E_q$ is nef.
(In fact, this is the class of a fiber of $S \to \mathbb{P}^1$.)
\end{enumerate}

\item (Continuation of previous item.)
$-K_S$ nef + genus formula  $\implies $ 
$C^2 \geq -2$ for every $C\subset S$ irreducible and rational.
This is an elliptic surface!

[We have taken a 1-dimensional linear system.
But why not take a higher-dimensional linear system?]

\item Fix $p_1,…,p_6$ in  $\mathbb{P}^2$ in general position
(not three in a line, not six in a conic).
Consider $f_1,f_2,f_3,f_4 \in \mathbb{C}[x_0,x_1,x_2]$ 
homogeneous of degree $3$ vanishing at $p_i$
and consider
\begin{align*}
\mathbb{P}^2 &\xymatrix{\ar@{-->}[r]&}\mathbb{P}^3 \\
p &\longmapsto (f_1(p):f_2(p):f_3(p):f_4(p)).
\end{align*}

\noindent
The resolved morphism $Bl_6\mathbb{P}^2 \to \mathbb{P}^3$ 
is an embedding with image of degree $3$
[a cubic surface on $\mathbb{P}^3$].
\end{enumerate}



\section{Elliptic surfaces}
\label{section-elliptic-surfaces}

\begin{definition}
\label{definition-elliptic-surface}
An {\it elliptic surface} 
is a surface $S$ together with $f:S\to C$ 
holomorphic 
(with connected fibers)
from $S$ to a smooth curve $C$ 
such that the generic fiber 
is a smooth curve of genus 1.

Moreover, we assume that there
do not exist $(-1)$-curves on fibers of $f$.
\end{definition}

Warning: I don't assume the existence
of a section
[a distinguished point on the fibers],
and I'm not saying that
all fibers are smooth.

\medskip\noindent
Singular fibers have been classified by Kodaira:
$I_0,I_1,(m)I_n,II, IV,IV^*=\tilde{E}_6,
III^*=\tilde{E}_7, II^*=\tilde{E}_8$ and $I_n$.

\begin{example}
\label{example-pencils-of-cubics}
$$
\left\{\substack{\text{pencils} \\ 
\text{of cubics}\\
\text{w/ smooth}\\
\text{member}}\right\}
\xymatrix{\ar@{<~>}[r]&}
\left\{\substack{\text{RES} \\ \text{w/ section}}\right\}.
$$
$$
\mathcal{P} \mapsto S_{\mathcal{P}}.
$$
 Question: given $\mathcal{P}$ and $S_{\mathcal{P}}$ 
how many singular fibers does $S_{\mathcal{P}}$ have?
\end{example}

\begin{lemma}
\label{lemma-discriminant-locus}
The discriminant locus in $\mathbb{P}^9$
[this $\mathbb{P}^9$ parametrizes cubics in $\mathbb{P}^2$ ]
has degree 12.
\end{lemma}

\noindent
Note that the Hodge diamond $S_{\mathcal{P}}$ 
is 
$$
\xymatrix@C=.5em@R=.5em{
& & 1\\
& 0 && 0\\
0 && 10 && 0\\
& 0 &&0\\
& & 1
}
$$
(Notice that adding up the center column gives 12.)

\begin{example}
\label{example-pencil-of-conics}
[We take a conic $C$ 
and a triple line $3L$ 
(it has to be triple so that it is a pencil of conics)
which is tangent to $C$ at an inflection point.
Then we blow up the point of intersection 9 times,
each time picking a point in the intersection
of the strict transform of the original cubic
and the triple line.]

Blowing up 9 times give
a surface $S_{\mathcal{P}}$ with
a $II^*$ fiber.
\end{example}

\medskip\noindent
Question: How is Kodaira's classification
obtained?

\begin{enumerate}
\item If $F=\sum n_iC_i$ is singular fiber,
then the intersection form restricts
to 
$\text{span}(\{C_i\})$.
This restriction is negative semi-definite
with a one-dimensional kernel spanned by $F$.

\item If $F$ is irreducible and $p_a(F)=1$,
then $F=$smooth,nodal or cuspidal.
 If $F$ is reducible, then each $C_i$ is
a $(-2)$-curve.
\end{enumerate}

\noindent
This two items combined imply that the
dual graph [vertices are curves,

two vertices joined if they intersect]
is a Dynkin diagram (ADE type)
[by a result of graph theory].




\section{The Weistrass model}
\label{section-weierstrass-model}

\noindent
Let $S \to C$ be an elliptic surface
with a section.
A choice of a section defines
a divisor of degree 1 on the generic fiber $S_\eta$,
i.e. a point $p$.

Looking at $H^0(S_\eta,np)$
for every $n$, we get:
\begin{align*}
H^0(p)&=\left<1\right>,\qquad  H^0(2p)=\left<1,x\right>,
\qquad H^0(3p)=\left<1,x,y\right>,\\
…,&\qquad   H^0(6p)=\left<1,x,y,x^2,xy,x^3,y^2\right>,
\end{align*}

\noindent
so there must exist a relation.
Up to further tricks,
we get to $y^2=x^3+ax+b$.

Making the construction global
gives the Weierstrass model.

\begin{itemize}
\item Any $S \to C$ is birational to a
Weierstrass fibration $W \to C$ 
(where $W$ can be singular, while $C$ is smooth)
where all fibers are irreducible and have $p_a=1$.
\end{itemize}

\noindent
Geometrically, $W$ is obtained from $S$ 
by contracting all fiber components that
do not meet the chosen section

[Once we show that we can do this,
it will suffice to study Weierstrass fibrations.]

[From now on $W$ is a Weierstrass fibration.]
Given $\pi_i:W \to C$ (with a section $\sigma$)
consider the fundamental line bundle
$\mathcal{L}=(\pi_*N_{\sigma/K})^\vee=(R^1\pi_*\mathcal{O}_W)^\vee$.
Then for all $n \geq 2$ 
we have a splitting
$\pi_*\mathcal{O}_{W}(n\sigma)\simeq \mathcal{O}_C \oplus \mathcal{L}^{-2}
\oplus … \oplus \mathcal{L}^{-n}$.
In particular, we have a section
$$
\pi^*\pi_*\mathcal{O}_W(3\sigma) 
\xymatrix{\ar@{->>}[r]&}\mathcal{O}_W(3\sigma)
$$
gives a map $f:W \to \mathbb{P}(\pi_*\mathcal{O}_W(3\sigma))$
that exhibits $W$ as a relative cubic in a $\mathbb{P}^2$-bundle
over $C$.

A global equation is given by
$$
Y^2Z=X^3+AXZ^2+BZ^3
$$
where $(Z,X,Y,A,B)$ are interpreted
as sections of 
$(\mathcal{O}_C,\mathcal{L}^2,\mathcal{L}^3,\mathcal{L}^4,\mathcal{L}^6)$.
$\sigma$ is given by $Z=X=0$.

\begin{remark}
\label{remark-singular-fibers}
The singular fibers of a minimal resolution of $W$ 
are determined by the order of vanishing of $A,B$ and
$\Delta=4A^3+27B^2$.
\end{remark}

\medskip\noindent
Consequences:
\begin{enumerate}
\item $\omega_W=\pi^*(\omega_C \otimes \mathcal{L})$ 
implies that $K_W^2=0$.
$e=R\cdot \text{deg} \mathcal{L}=$
number of singular fibers.

\item $\kappa(W) \leq  1$.

\item $\text{deg} \mathcal{L}=\chi(W)$, $e=R\cdot \text{deg} \mathcal{L}=$
number of singular fibers.

\item If $C=\mathbb{P}^1$,
then $\mathcal{L}=\mathcal{O}_{\mathbb{P}^1}(d)$ 
for some $d$.

$$
\begin{cases}
d=1\xymatrix{\ar@{<~>}[r]&}&\text{(general) RES w/section} \\
d=2\xymatrix{\ar@{<~>}[r]&}&\text{ K3s, }K_W=0\\
3\geq 3\xymatrix{\ar@{<~>}[r]&}& \kappa=1\text{ and }K_W=\alpha F
\text{for some }\alpha>0.
\end{cases}
$$
\end{enumerate}

\medskip\noindent
Invariants. $\pi:W \to C$.
Set $g=g(C)$.

\begin{center}\begin{tabular}{c|c c}
& $q$ & $p_g$\\
\hline
$W$ not a product &  $g$ & $ g+ \text{deg}\mathcal{L}-1$ \\
product& $g+1$ &  $g+\text{deg}\mathcal{L}$.
\end{tabular}\end{center}

\noindent
We can also compute the pluri-genera $P_n$ 
and deduce:
\begin{itemize}
\item $g=0$: done.

\item $g=1$:
 $$
\begin{cases}
\mathcal{L}=\mathcal{O}_C\qquad &E\times E \\
\mathcal{L}\text{ is torsion of order 2,3,4 or 6 hyp}\\
\text{deg}(\mathcal{L})\geq 1\qquad &\kappa=1
\end{cases}
$$

\item $g\geq 2$: $\kappa=1$.
\end{itemize}

\begin{proposition}
\label{proposition-minimal-surface}
Let $S$ be a minimal surface with $\kappa=1$.
Then
\begin{enumerate}
\item $K^2_S=0$ and
\item there exists $f:S \xymatrix{\ar@{->>}[r]&} C$ elliptic fibration.
\end{enumerate}
\end{proposition}

\begin{proof}
\begin{itemize}
\item $S$ contains a curve
$E= \sum n_iC_i$
($n_i \in \{ 0,1\}, \text{supp}(E)$ is connected)
with $E \overset{num}{\not \sim}0$ 
and $E^2=K_S E=0$.

In fact $K_S \overset{num}{\sim} rE$ for some $r \in \mathbb{Q}$.

\item Some multiple of $E$ moves in an elliptic pencil.
Some multiple of $E$ is the movable part of $\tilde{r}K_S$ 
for some $\tilde{r} \in \mathbb{Q}$.
\end{itemize}
\end{proof}

\section{Examples}
\label{section-examples}

\begin{itemize}
\item (The Fermat quartic.)
Consider $S \subset \mathbb{P}^3$ 
given by $x_0^4+x_1^4+x_2^4+x_3^4=0$ 
and for each $(\lambda:\mu) \in \mathbb{P}^1$,
let
$$
C_{(\lambda:\mu)}
=\begin{cases}
\lambda(x_0^2+\zeta^2x_1^2-\mu(x_2^2-\zeta^2x_3^2=0\\
\mu(x_0^2-\zeta^2x_1)+\lambda(x_2^2+\zeta^2x_3^2)=0
\end{cases}
$$
where $\zeta \in \mathcal{M}_8$
(an eight root of unity), primitive.
These curves live in $S$ and
\begin{align*}
\pi: S &\longrightarrow \mathbb{P}^1 \\
(x_0:x_1:x_2:x_3) &\longmapsto 
(x_0^2-\zeta^2x_1^2:-x_2^2-\zeta x_3^2)
\end{align*}

\noindent
is an elliptic surface.

\item Consider a smooth plane quartic $Q$
and choose a point $p \not\in Q$.
$$
\xymatrix{
S \ar[d]^{Bl_{q_1,q_2}Y \text{ where
$\pi^{-1}(p)=\{q_1,q_2\}$}}\\
Y \simeq Bl_7\mathbb{P}^2\ar[d]_\pi^{2:1}\\
\mathbb{P}^2 \supset Q.
}
$$
where $\pi^{-1}(\ell)$ has $g=1$.

[This produces a rational elliptic surface 
with a section]
\end{itemize}

\noindent
Notice that there exist elliptic surfaces with multiple fibers
(hence without a section).

\begin{example}
\label{example-multiple-fibers}
Start with three lines in the plane
intersecting pairwise, and a conic intersecting tangentially
once every line [Fano plane].
We blow up these three points and obtain
a type $2I_3$ fiber.

In general, we take pencils of sextics
with nodes at 9 points. These are in correspondence
with Halpen surfaces.
\end{example}





\section{Halphen surfaces}
\label{section-halphen-surfaces}

\noindent
[These are a class of rational surfaces.]

\begin{proposition}
\label{proposition-res}
Let $\pi:S \to C$ be an elliptic surface with $S$ rational.
Then
\begin{enumerate}
\item $C=\mathbb{P}^1$, $b_2(S)=10$.

\item $\pi$ has at most one multiple fibre $mF$
(necessarily type  $m I_n$
for some $n$ and some $m$.

\item either $\pi$ has no multiple fiber
and $-K_S \sim F$ (any fiber),
or $\pi$ has one muliple fiber
of multiplicity $m$ and $-m K_S \sim S_p$

\item if $\pi$ has no multiple fibers, there exists a section.
\end{enumerate}
\end{proposition}

\begin{proof}[Proof of 2. and 3.]
Canonical bundle formula:
\begin{equation}
\label{equation-canonical-bundle-formula}
\omega_S=\pi^* (\omega_{\mathbb{P}^1}\otimes\mathcal{O}_{\mathbb{P}^1}(1))
\otimes \mathcal{O}_S\left(
\sum_p(m(p)-1)S_p)\right).
\end{equation}
Then, given any smooth fiber $F$,
for any $n \in \mathbb{N}$,
we have $n K_S \sim n\left(\sum_p\frac{
m(p)-1}{m(p)}-1\right)^\alpha F$.

Note that $\alpha<0$ because no multiple
of $K_S$ can be effective.
Thus, $m(p)=1$ except for at most $1$ at $p$.

Why is \ref{equation-canonical-bundle-formula} true?
In general, for any relatively minimal
elliptic surface $\pi:S \to C$
with multiple fibers
$m_1F_1,…, m_k F_k$:
$$
\omega_S
=\pi^* (\omega_C \otimes
\underbrace{R^1\pi_*\mathcal{O}_S}_{\text{deg}\chi})^\vee)
\otimes \mathcal{O}_S\left(\sum m(p)-1S_p\right).
$$
Why? Because $\pi_*\omega_S$ is locally free,
thus $\omega_S=\pi^*(\omega_C \otimes \mathcal{L})\otimes\mathcal{O}_S(D)$
where $D$ is the zero divisor
of $\pi^*\pi_*\omega_S\to \omega_S$.
Argue that $D=\sum n_iF_i$
with $n_i< m_i$ using adjunction:
$$
\mathcal{O}_{F_i}\simeq \omega_{F_i}
=\omega_S \otimes\underbrace{\mathcal{O}_{F_i}(F_i)}_{\text{torsion}} 
=\mathcal{O}_{F_i}(F_i)^{\otimes (n_i+1)}
$$
thus $n_i+1$ has to divide $m_i$.
\end{proof}

\begin{remark}
\label{remark-rational-elliptic-surface}
Let $\pi:S \to \mathbb{P}^1$ be a rational elliptic surface (RES)
and note that $K_S^2=0$.
So, $S$ is not minimal and there exists a  $(-1)$-curve.
By Proposition \ref{proposition-res},
there are two possibilities:
\begin{enumerate}
\item There are no multiple fibers, $M.S_n=1$.
\label{item-no-multiple-fibers}

\item There exists a multiple fiber $mF$ 
\label{item-halphen}
and $MS_n=m$,
\end{enumerate}
\end{remark}

\begin{definition}
\label{definition-halphen-surface}
A RES as in Item \ref{item-halphen}
is called a {\it Halpen surface} of index $m$.
Alternatively, a rational surface  $S$ is
a {\it Halphen surface} if there exists $m$ 
such that $|-m K_S|$ is one dimensionsional
and has no fixed component and no base points.
The smallest such $m$ is the {\it index}.
(We include $m=1 \iff$ there exists a section.)
\end{definition}

\begin{proposition}
\label{proposition-halphen}
If $\pi:S \to \mathbb{P}^1$ 
is a Halphen surface
of index $m$, there exists $f:S \to \mathbb{P}^2$ 
birational such tat $\pi \circ f^{-1}$ is a {\it Halpen pencil}
of index $m$
(i.e. a pencil of plane curves
with degree $3M$ with 9 base points,
each of multiplicity $m$).
Conversely, the resolution
of any such pencil gives a Halpen surface.
\end{proposition}

\begin{example}
\label{example-halphen}
$Q:y^2+xz=0$,
$L:y=0$,
$C:(y^2+xz)(\alpha y+z)+\beta yx^2$,
$\alpha,\beta \neq 0$.

The pencil $\lambda(4Q+L)+\mu(3C)=0$
is a Halphen pencil of index 3.

[The intersection of $C$, $Q$ and $L$ consists of two points,
$p_1=(0:0:1)$ and $p_2=(1:0:0)$.]
Blow up at $p_1$ seven times
and at $p_2$ two times.
[We obtain one the exceptional Dynkin diagram fibers.]
\end{example}

\begin{example}
\label{example-halpen2}
Consider any smooth cubic $C$.
Take $p_1,…,p_9 \in C$ such that
$p_1 \oplus … \oplus p_9$ is $m$-torsion.
Then there exists a curve $D$ of degree $3m$ passing
though $p_1,…,p_9$ with multiplicity $m$,
thus $\lambda(D)+\mu(mC)=0$ is a Halpen pencil.
\end{example}

\section{Singular fibers}
\label{section-singular-fibers}

Recall that
$$
\left\{\substack{\text{Halphen pencils} \\ 
\text{$\lambda C_{3m}+\mu(mC)=0$}}\right\}
\xymatrix{\ar@{<~>}[r]&}
\left\{\substack{
\text{Halphen surfaces} \\ 
\pi:S \xrightarrow{|-m K_S|}\mathbb{P}^1}\right\}
$$
$$
C_{3m}\xymatrix{\ar@{<~>}[r]&} F\in |-m K_S|.
$$

\begin{proposition}[Z]
\label{proposition-singular-fibers}
If $F$ is of type $II^*$, $III^*$ or  $IV^*$,
then $C_{3m}$ is non-reduced.
\end{proposition}

\begin{proof}
$R\subset S$ rational smooth $\implies $ $R^2 \geq  -2$,
$\implies $ exceptional curves in $S$ can only appear
in disjoint chains.
Now, if $F$ is as in the statement,
then at least one curve in [picture of fiber]
is not exceptional.
Here are the fibers: 
\end{proof}

\begin{proposition}[Z]
\label{proposition-halpen}
Let $S \to \mathbb{P}^1$ be a Halphen surface of index 2
and assume that $\pi$ contains a fiber $F$ of type $II^*$.
Then there exists $\varphi:S \to \mathbb{P}^2$ such that
the sextic $\varphi(F)$ is one of the following:
[Picture of conic of multiplicity 3,
line of multiplicity 3 intersecting cubic,
two lines of multiplicity 3,
quadric and line of multiplicity 4,
a line and a line with multiplicity 5].
\end{proposition}

\begin{example}
\label{example-halphen3}
$C:y^2z=x(x-2)(x-\alpha z)$, $\alpha \neq 0,1$.
$L:z=0$,
$\tilde{L}:x=0$,
$\mathcal{P}:\lambda(L+5\tilde{L})+\mu(2C)=0$.

[A point of intersection is $p_1=(0:1:0)$,
where the line is tangent at the inflection point of $C$.
Another point of intersection is $p_2=(0:0:1)$.]
Blow up $p_1$ 4 times and $p_2$ 5 times.
\end{example}

\section{Constructions}
\label{section-construction}
\begin{enumerate}
\item 
$$
\xymatrix{
X\ar[d]^{m:1}\\
S \ar[d]^{|-m K_S|\ni F \text{ smooth}}\\
\mathbb{P}^1
}
$$
For  $m=2$,  $X$ is a K3. For $m\geq  3$,
$\kappa(X)=1$ and contains a $(-m)$ $P^1$ section.


\item Type $II$ degenerations of K3 surfaces
and degenerations of Enriques surfaces
(see Alexeev and Engel).

\item It $\pi:S \to \mathbb{P}^1$ is a general Halphen surface
of index 2 and we blow up we get a Coble surface.

\item  The Halpen transform
$$
\xymatrix{
S \ar[d]^{|-m K_S|}\\ \mathbb{P}^1
}
$$
$|F+M|$ curves of $g=1$ through $q$.
Contract $M$ and blow up $q$,
$\xymatrix{\ar@{~>}[r]&}$ we obtain a RES with section.

\item (Halpne.)
Given $p_1,…,p_8 \in \mathbb{P}^2$,
the locus of the 9th base point of a Halpen pencil of index $m$ 
is a curve of degree $3\psi(m)$
having ordinary multiple points at $p_1,…,p_8$ 
of multiplicity $\psi(m)$,
where
 $$
\psi(m)=m^2\left(1-\frac{1}{q_2^2}\right) 
\left(1-\frac{1}{q_2^2}\right)…
$$
where $m=q_1^{r_1}q_2^{r_2}…$ 
prime factorization. $m=2 \implies \psi(m)=3$.

\end{enumerate}


\section{K3 surfaces}
\label{section-k3}

\begin{definition}
\label{definition-k3}
A surface $S$ is a {\it K3} if
$\Omega^2=\mathcal{O}_S$ and $h^1(S,\mathcal{O}_S)=0$.
\end{definition}

\noindent
So, $\omega_S=\mathcal{O}_S$.
Also $\Omega^2=\mathcal{O}_S$ means that there exists
a global nowhere zero holomorphic 2-form $\sigma$.

\begin{example}
\label{example-k3}
\begin{enumerate}
\item $S \subset \mathbb{P}^3$ smooth quartic.
$$
\xymatrix{
0\ar[r]
&\mathcal{O}_{\mathbb{P}^3}(-4)\ar[r]
&\mathcal{O}_{\mathbb{P}^3}\ar[r]
&\mathcal{O}_S\ar[r]
&0
}
$$
The long exact sequence in cohomology
gives $h^1(S,\mathcal{O}_S)=0$.
Adjunction gives
$\omega_S=\omega_{\mathbb{P}^3}\otimes\mathcal{O}_{\mathbb{P}^3}(4)|_S
=\mathcal{O}_S$.

\item $\pi:S \xrightarrow{2:1}\mathbb{P}^2 \supset B$
smooth sextic.
$\pi_*\mathcal{O}_S
=\mathcal{O}_{\mathbb{P}^2}\oplus\mathcal{O}_{\mathbb{P}^2}(-3)$ 
gives $h^1(S,\mathcal{O}_S)=0$.
(If $\mathcal{L}=\mathcal{O}_{\mathbb{P}^2}(-3)$,
$B$ is a section of $\mathcal{L}^{-2}$.)
Since $\omega_S=\pi^* (\omega_{\mathbb{P}^2}\otimes\mathcal{L}^{-1}$.
This implies that $\omega_S \simeq \mathcal{O}_S$.
\begin{remark}
\label{remark-ade-singularities-on-k3}
We can allow ADE singularities on $B$ and take a minimal resolution
$$
\xymatrix{
\tilde{S}\ar[d]_{2:1}&S\text{ K3 surface}\ar[d]^{2:1}
\ar[l]_{\substack{ \text{min}\\\text{res} }}\\
\mathbb{P}^2\ar@{-->}[d]^{\substack{B \\ \text{sextic}\\
\text{w/ 9 noded}}}& \text{ Halphen index 2}\ar[d]^{|-2K_Y|\ni F}\\
\mathbb{P}^1\ar[r]&\mathbb{P}^1
}
$$
\end{remark}

\item $S$ a K3 surface with a fixed point free involution $i$ 
then $S/i$ is an {\it Enriques surface} [a friend of K3],
$p_g=q=0,2K=0$.

\item (Kummer surfaces.) $A$ abelian surface (e.g. $E \times E^1$)
$a \mapsto  -a$ has 16 fixed points.
$Kum(A):=\widetilde{A/(-1)}$ is a K3 surface.
\end{enumerate}
\end{example}

\noindent
The Hodge diamond of a K3:
$$
\xymatrix@C=.5em@R=.5em{
& & 1\\
& 0 && 0\\
1 && 20 && 1\\
& 0 && 0\\
& & 1
}
$$
[The $H^2$ lattice is]
$$
H^2(S,\mathbb{Z})=E_8(-1)^{\oplus 2} \oplus U^{\oplus 3},
$$
an even unimodular lattice of signature 
$(\underbrace{3}_{>0},\underbrace{19}_{<0})$.

\section{Curves on a K3 surface}
\label{section-curves-on-k3}

Let $S$ be a K3 surface and $C \subset S$ a curve.
By adjunction formula, $C^2=2p_a(C)-2$, so $C^2 \geq -2$.

In particular, if $C$ is smooth and $C^2=-2$,
$C=\mathbb{P}^1$.

Assume that $C$ is smooth and irreducible, and set $g=g(C)$,
$\mathcal{L}=\mathcal{O}_S(C)$.

Then $\mathcal{L}^2=2g-2$, $h^0(S,\mathcal{L})=g+1$.
Moreover, if $g \geq 1$,
then $|C|$ is base point free and the induced morphism
$S \to \mathbb{P}^9$ restricts to the canonical map on $C$.

\begin{definition}
\label{definition-polarized-k3}
A {\it polarized K3 surface} of degree $2d$ is a pair
$(S,\mathcal{L})$ with $S$ a K3 and  $\mathcal{L}$ an
ample (ind in Pic) line bundle of degree $\mathcal{L}^2=2d$
\end{definition}

\begin{proposition}
\label{proposition-polarized-k3}
For any $g \geq 3$ there exists a polarized K3
of degree $2g-2$ in $\mathbb{P}^9$.
\end{proposition}

\begin{example}
\label{example-polarized-k3}
\begin{itemize}
\item $2d=4$. Smooth quartic in $\mathbb{P}^3$ 
or a double cover of $\mathbb{P}^1 \times \mathbb{P}^1$ branched
aloong a $(4,4)$ curve.

\item $2d=2$.  $S \xrightarrow{2:1}\mathbb{P}^2 \supset$ sextic.
\end{itemize}
\end{example}

\section{Elliptic K3s}
\label{section-elliptic-k3}

\begin{enumerate}
\item 
$$
\xymatrix{
&Kum(E \times E^1)\ar[dl]\ar[dr]\\
\mathbb{P}^1\simeq E/(-1)& & \mathbb{P}^1 \simeq E^1/(-1)
}
$$

\item
$$
\xymatrix{
S \ar[r]^{\substack{\text{min} \\ \text{res}}}
&\tilde{S} \ar[r]^{2:1}
&\mathbb{P}^1 \supset \substack{\text{sextic} \\ \text{w/ node}}.
}
$$
The elliptic pencil comes from the pencil of lines in $\mathbb{P}^1$ 
through the node.

\item 
$$
\xymatrix{
S \text{ elliptic}\ar[d]^{2:1}\\
\underbrace{F_1 \cup F_2}_{\text{smooth}}\quad 
X \text{ a RES w/ section}\ar[d]\\
\mathbb{P}^1
}
$$
\end{enumerate}

\begin{proposition}
\label{proposition-elliptic-k3}
Let $S$ be a K3 surface.
\begin{enumerate}
\item $S$ is elliptic $\iff \exists  L \in \text{Pic}(S)$ with $L^2=0$.

\item $S$ is elliptic with a section $\iff U \xymatrix{\ar@{^{(}->}[r]^{
\text{primitive}}&} \text{NS}(S)$ [where $U$ is the hyperbolic lattice].

\item (If the Picard rank is big enough, the surface will be elliptic.)
If $\rho(S) \geq 5$, then $S$ is elliptic
$\rho(S) \geq  12 \implies \exists $ a section.
\end{enumerate}
\end{proposition}

To visualize the first point of the proposition,
consdier the following example.

\begin{example}
\label{example-smooth-quartic-elliptic}
A smooth quartic in $\mathbb{P}^3$ is elliptic iff it contains a line.
\end{example}

\begin{remark}
\label{remark-elliptic-k3}
\begin{itemize}
\item A K3 may admit more than one elliptic fibration.

\item Since the base is always $\mathbb{P}^1$
[as you remember from the examples],
there will always be a singular fiber.
If $\underbrace{S}_{K3} \to C$ with $g(C) > 0$
then  $h^0(\Omega_C)>0$ 
but $H^0(\Omega_C) \xymatrix{\ar@{^{(}->}[r]&} H^0(\Omega_S)$.

\item Any 
$\underbrace{S}_{K3} \xymatrix{\ar@{->>}[r]^{\text{flat}}&}\mathbb{P}^1$ 
is an elliptic fibration.
\end{itemize}
\end{remark}

\section{The Weierstrass model}
\label{section-weierstrass-model}

\noindent
For an elliptic K3 $S \to \mathbb{P}^1$ with section,
$R^1\pi_*\mathcal{O}_S = \mathcal{O}_{\mathbb{P}^1}(-2)$.
So, a Weierstrass model is given by
$$
y^1=x^3 + A(t)x + B(t),\qquad t \in \mathbb{P}^1
$$
with $\text{deg}A=8,\text{deg}B=12$.

\medskip\noindent
The volume form can be written locally as
$$
\frac{dx\wedge dt}{2y}.
$$
\section{Automorphism}
\label{section-automorphisms}

\noindent
[Automorphisms will act in differential forms, cohomology, etc.
A fundamental question is if an automorphism will
preserve the volume form or not.]

$$
\xymatrix{
&  \text{symplectic }\iff \sigma \text{ is preserved}\\
\text{2 types}\ar@{-}[ur]\ar@{-}[dr]\\
&  \text{non- symplectic otherwise}
}
$$
(Purely non-symplectic if it acts as multiplication by a 
primitive $n-$-th root of 1.)

\noindent
We restrict to finite order.

\begin{proposition}
\label{proposition-restrictions-on-orders}
Let $S$ be a K3 surface
with volume form $\sigma$.
\begin{enumerate}
\item If $f \in \text{Aut}(S)$ acts as
$f^*\sigma=\underbrace{\zeta_s}_{\text{primitive}}\sigma$,
then $\varphi(n) \leq  20$ and all $n \neq  20$ satisfying
this appear.

If $n=2$, $\text{Fix}(f)$ contains only curves
(no isolated points) or $\emptyset$.

If $n \geq 3$, $\text{Fix}(f)
=C_g \coprod R_1 \coprod … \coprod R_k \coprod \{p_1,…,p_m\}$,
$g \geq 0$. (Nikulin, Sarti+Artebani, Hoffmann Brandhorst.)

\item If $f^*\sigma=\sigma$,
then $\text{Fix}(f)=\{p_1,…,p_k\}$ with $1 \leq k \leq 8$
and $\text{ord}(f) \leq 8$.
\end{enumerate}
\end{proposition}

\begin{center}
\begin{tabular}{c|c c c c c c c}
n & 2&3&4&5&6&7&8\\
\hline
$\text{Fix}(f)$&8&6&4&4&2&3&2\\
$\rho(S)$ &9&13&15&17&17&19&19
\end{tabular}
\end{center}

\noindent
(See Salgado, Garbagnati. Comparin-Prieto-Troncoso-Montero (on involutions).)

\begin{example}
\label{example-automorphisms}
Consider the elliptic K3 given by
[the Weierstrass model …]
$$
y^2=x^3+t^2x+t^{10}
$$
[we have written in this in a simpler way, check degrees]
and the automorphism
$$
f:(x,y,t)\mapsto (\zeta_7x,\zeta_7^5y,-\zeta_7t)
$$
(order 14).

\noindent
Note: 
\begin{align*}
A(t,s)&=t^2s^6\\
B(t,s)&=t^{10}s^2\\
\Delta(t,s)&=4A^3+27B^2
=\text{const. }s^4t^6(s^{14}-t^{14}).
\end{align*}

\begin{itemize}
\item At $t=0$,  $a=2,b=10,\delta=6$. $I_0^*$.

\item At $s=0$, $a=6, b=2, \delta=4$. $IV$.

\item At each root of $s^{14}-t^{14}$,
$a=b=0,\delta=1$, $I_1$.
\end{itemize}

\noindent
[Pictures of the fibers and how the 
automorphisms act on them - fixed components etc.]
\end{example}



















\end{document}
